Report - Latest Rodent Data

### 2019-02-19 index.Rmd
tidying the website Rmds
this will not pass a full build (including website rendering) until v 0.2.3 of portalcasting is merged in and a new docker image is made

Welcome to Portal Forecasting! This is a website run by the [Weecology](http://weecology.org/) group, a group of interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/). On this website, you'll find information about our ongoing efforts to forecast ecological systems from time series. Specifically, we are using a times series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology. Enjoy!

@@

This is the forecast for next month's sampling of rodents at Portal. The black line indicates the historic total rodent abundance. The blue line indicates the forecasted amount over the next 12 months, along with confidence intervals. This forecast is from the Ensemble model.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
obs_data = portalr::abundance(clean=FALSE, level='treatment', plots="longterm") %>%
  dplyr::filter(treatment=="control")
obs_data$total = rowSums(dplyr::select(obs_data, -c(period, treatment)))
new_moons = read.csv('PortalData/Rodents/moon_dates.csv') %>%
                   dplyr::mutate(censusdate = as.Date(censusdate),
                          year = lubridate::year(newmoondate),
                          month = lubridate::month(newmoondate))

obs_data_newmoon = dplyr::inner_join(obs_data, new_moons, by = c("period" = "period"))

for_data = compile_forecasts() %>%
  dplyr::filter(level == 'Controls', model == 'Ensemble', date == max(date))

# get dates for future new moons for plotting
future_moons = new_moons %>% portalr::get_future_moons() %>%    
                         dplyr::select(c("newmoonnumber","newmoondate"))
for_data = dplyr::inner_join(for_data, future_moons, by = c("newmoonnumber" = "newmoonnumber"))

forecast_viz(obs_data = obs_data_newmoon,
             obs_date_col_name = "censusdate",
             obs_val_col_name = "total",
             for_data = for_data,
             for_date_col_name = "newmoondate",
             for_val_col_name = "estimate",
             for_model_name = "Ensemble",
             for_lowerpi_col_name = "LowerPI",
             for_upperpi_col_name = "UpperPI",
             start_newmoon = 300,
             ylabel = 'Total Abundance')
```

@@

```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}

sp_predictions = get_sp_predicts(for_data, 'Controls', 
            lead_time = length(unique(for_data$newmoonnumber)) - 11)
sp_predict = plot_species_forecast(sp_predictions,
            title=paste0(sp_predictions$forecast_date[2], ": Control plots"))

plot(sp_predict)
```

@@

This figures are similar to the primary Total Abundance forecast above, but for the abundance of three common rodent species. 
```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE}
most_abund_sp = sp_predictions %>% 
  dplyr::filter(species != 'total') %>% 
  dplyr::arrange(desc(estimate)) %>% 
  head(3) %>% 
  dplyr::select(species)

# load in rodent species table to get scientific names to display on plots
species_table = read.csv('PortalData/Rodents/Portal_rodent_species.csv',
                          stringsAsFactors = F,na.strings = '')
most_abund_sp_names = dplyr::filter(species_table,speciescode %in% most_abund_sp$species) %>%      
                      dplyr::select(speciescode,scientificname)

for (n in seq(dim(most_abund_sp_names)[1])) {
  species_forecast = forecast_viz(obs_data = obs_data_newmoon,
               obs_date_col_name = "censusdate",
               obs_val_col_name = most_abund_sp_names$speciescode[n],
               for_data = for_data,
               for_date_col_name = "newmoondate",
               for_val_col_name = "estimate",
               for_model_name = "Ensemble",
               for_lowerpi_col_name = "LowerPI",
               for_upperpi_col_name = "UpperPI",
               start_newmoon = 300,
               ylabel = most_abund_sp_names$scientificname[n])
  plot(species_forecast)
}
```

>>>>

Welcome to Portal Forecasting! This is a website run by the [Weecology](http://weecology.org/) group, interdisciplinary ecologists at the [University of Florida](http://www.wec.ufl.edu/). On this website, you'll find information about our ongoing efforts to forecast a time series of rodent abundances from [The Portal project](http://portal.weecology.org/), a long-term experimental monitoring project in desert ecology. Enjoy!

@@

This is the forecast for next year's sampling of rodents in the control plots at Portal. The black line indicates the historic total rodent abundance. The blue line indicates the Ensemble model's forecast for total rodent abundances over the next 12 months, along with 90% confidence intervals.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
plot_cast()
```

@@

```{r speciesforecasts, echo=FALSE, message=FALSE, warning=FALSE}
plot_species_casts()
```

@@

These figures are similar to the primary Total Abundance Forecast above, but for the abundance of three common rodent species. 
```{r highabund_ts_forecasts, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
most_ab <- select_most_ab_spp()
for(i in 1:3){
  plot_cast(species = most_ab[i])
}
```

### 2019-02-19 report.Rmd
tidying the website Rmds
this will not pass a full build (including website rendering) until v 0.2.3 of portalcasting is merged in and a new docker image is made

Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
```{r, echo=FALSE, message=FALSE, warning=FALSE}

# get most recent rodent data
observed = portalr::abundance(clean=FALSE, shape='flat', level='treatment', plots='longterm', time = 'newmoon') %>%
            dplyr::filter(newmoonnumber == max(newmoonnumber), treatment=='control')

# get new moon number associated with this period
most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
                   dplyr::pull(censusdate)

# get most recent forecast prior to data collection
files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)

data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')

ensemble = dplyr::filter(data, level == 'Controls', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
sp_predictions = get_sp_predicts(ensemble, 'Controls', lead_time = 1)

joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
joined_data[is.na(joined_data)] = 0
joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)

sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": Control plots"))  
sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
             color = "blue")
plot(sp_predict)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# get most recent rodent data
observed = portalr::abundance(clean=FALSE, shape='flat', time = 'newmoon') %>%
            dplyr::filter(newmoonnumber == max(newmoonnumber))

# get new moon number associated with this period
most_recent_date = read.csv('PortalData/Rodents/moon_dates.csv') %>%
                   dplyr::mutate(censusdate = as.Date(censusdate,format='%Y-%m-%d')) %>%
                   dplyr::filter(newmoonnumber==unique(observed$newmoonnumber)) %>% 
                   dplyr::pull(censusdate)

# get most recent forecast prior to data collection
files = data.frame(names = list.files(path = './predictions', pattern = '*forecasts.csv'))
files$filedates = as.Date(substr(files$names,1,10),format='%Y-%m-%d')
latest_forecast = dplyr::filter(files,filedates<most_recent_date) %>% tail(1)

data = read.csv(paste('predictions/',latest_forecast$names,sep=''),na.strings = '')

ensemble = dplyr::filter(data, level == 'All', model == 'Ensemble', newmoonnumber==unique(observed$newmoonnumber))
sp_predictions = get_sp_predicts(ensemble, 'All', lead_time = 1)

joined_data = dplyr::left_join(sp_predictions, observed, by = c('species','newmoonnumber'))
joined_data[is.na(joined_data)] = 0
joined_data[joined_data$species=='total','abundance'] = sum(joined_data$abundance,na.rm=T)

sp_predict = plot_species_forecast(sp_predictions,title=paste0(sp_predictions$forecast_date[2], ": All plots"))  
sp_predict = sp_predict + ggplot2::geom_point(data = joined_data, mapping = ggplot2::aes(x = abundance, y = species),
             color = "blue")
plot(sp_predict)
```

>>>>

Here's the most recent actual data (blue) compared to the forecast performed right before data collection (black, with error bars)
```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_species_casts(with_census = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
plot_species_casts(with_census = TRUE, level = "All")
```
